DynaProg for Scala

### 2013-01-05 docs/report/implementation.tex
Nits.

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Code generation}
The code generation step produces multiple outputs that are tightly bound to each other. Besides the Scala wrapper (a simple JNI interface), in the C/CUDA code generated we distinguish:\ol
\item JNI input and output conversion functions
\item Host memory management and scheduling of CUDA kernels
\item CUDA matrix computation, which can be further decomposed into matrix scheduling (loops) and (matrix cell) computation. 
\item CUDA backtrack collection kernel
\ole

@@

% ----------------------------------------------
\subsubsection{Parsers code generation}
Parsers generation is independent of user-defined function generation (see \ref{user_fun_gen}). Tabulation inner parsers are first wrapped in additional aggregation (by $h$, hereby ensuring they produce at most one result) and normalized (according to \ref{normalization}); code generation then occurs recursively, producing a list of loops and conditions, and body (possibly with a hoisted part). Additionally, position variables are maintained and sub-rule index and concatenation indices are propagated. We give an overview of each parser transformation:\ul
\item \textbf{Terminal:} provides its own C code, which correspond usually to the input element value, its position or the position of the matching range.
\item \textbf{Tabulate:} is a simple value load, possibly wrapped into a validity check. Useless validity verification can be removed by marking the tabulation as <<always valid>>.
\item \textbf{Aggregate:} corresponds to an intermediate (value,backtrack) pair where inner parsers write their result; outermost aggregation is written back to corresponding (cost, backtrack) matrices. Validity information, and concatenation indices are propagated within backtrack. To preserve a correct semantic, inner aggregations body is hoisted outside loops and condition checks of the enclosing parser.
\item \textbf{Or:} since parsers are normalized and operate on a single aggregation result, it suffice to emit sequentially code of alternatives.
\item \textbf{Map:} wraps its argument into a the user-defined function call
\item \textbf{Filter:} wraps its body into user-defined condition check
\item \textbf{Concat:} fixed size concatenation are wrapped in simple conditions; moving concatenations are wrapped in a {\tt for} loop. The loops and conditions are further simplified to reduce range and remove useless conditions before actual code is emitted.
\ule

@@

% ----------------------------------------------
\subsubsection{Backtracking on the GPU}
{\center\color{red} XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\\ CONTINUE HERE\\ 
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX}

XXX: explain how it is done, how to get list reversal for free, explain lattice of elements with construction order.

\begin{verbatim}
__global__ void gpu_backtrack(trace_t* trace, unsigned* size, back_t* back, int i0, int j0) {
  const unsigned trace_len[2] = {1,1};
  trace_t *rd=trace, *wr=trace; *size=0;
  #define PUSH_BACK(I,J,RULE) { wr->i=I; wr->j=J; wr->rule=RULE; ++wr; ++(*size); }
  PUSH_BACK(i0,j0,0);
  for(;rd<wr;++rd) {
    bt1* bt;
    switch (rd->rule) {
      case 0: bt=(bt1*)&back[idx(rd->i,rd->j)].M; break;
      case 1: bt=(bt1*)&back[idx(rd->i,rd->j)].M; break;
    }
    rd->rule=bt->rule;
    for (int i=0,l=trace_len[rd->rule]; i<l; ++i) rd->pos[i]=bt->pos[i];
    switch (rd->rule) {
      case 0: break;
      case 1: PUSH_BACK(rd->i,rd->pos[0],0); PUSH_BACK(rd->pos[0],rd->j,0); break;
    }
  }
}
\end{verbatim}

% ----------------------------------------------
\subsubsection{User functions generation} \label{user_fun_gen}
How LMS is used to generate function
Possibly write that we can do the same with macros
Why cannot use LMS for everything
Where and how is LMS useful

\subsection{Runtime execution engine}
XXX: compiler for C and for Scala

\subsection{LibRNA (?)}
The project development was scheduled in \ul
\item Problems analysis
\item Ad-hoc implementation (CUDA)
\item Porting ADP parsers
\item Improving parsers / adding functionalities
\item Integrating CUDA generation into parsers
\ule

>>>>

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Code generation} \label{codegen}
The code generation step produces multiple outputs that are tightly bound to each other. Besides the Scala wrapper (a simple JNI interface), in the C/CUDA code generated we distinguish:\ol
\item JNI input and output conversion functions
\item Host memory management and scheduling of CUDA kernels
\item CUDA matrix computation, which can be further decomposed into matrix scheduling (loops) and (matrix cell) computation. 
\item CUDA backtrack collection kernel
\ole

@@

% ----------------------------------------------
\subsubsection{Parsers code generation}
Parsers generation is independent of user-defined function generation (see \ref{user_fun_gen}). Tabulation inner parsers are first wrapped in additional aggregation (by $h$, hereby ensuring they produce at most one result) and normalized (according to \ref{normalization}); code generation then occurs recursively, producing a list of loops and conditions, and body (possibly with a hoisted part). Additionally, position variables are maintained and subrule index and concatenation indices are propagated. We give an overview of each parser transformation:\ul
\item \textbf{Terminal:} provides its own C code, which correspond usually to the input element value, its position or the position of the matching range.
\item \textbf{Tabulate:} is a simple value load, possibly wrapped into a validity check. Useless validity verification can be removed by marking the tabulation as <<always valid>>.
\item \textbf{Aggregate:} corresponds to an intermediate (value,backtrack) pair where inner parsers write their result; outermost aggregation is written back to corresponding (cost, backtrack) matrices. Validity information, and concatenation indices are propagated within backtrack. To preserve a correct semantic, inner aggregations body is hoisted outside loops and condition checks of the enclosing parser.
\item \textbf{Or:} since parsers are normalized and operate on a single aggregation result, it suffice to emit sequentially code of alternatives.
\item \textbf{Map:} wraps its argument into a the user-defined function call
\item \textbf{Filter:} wraps its body into user-defined condition check
\item \textbf{Concat:} fixed size concatenation are wrapped in simple conditions; moving concatenations are wrapped in a {\tt for} loop. The loops and conditions are further simplified to reduce range and remove useless conditions before actual code is emitted.
\ule

@@

% ----------------------------------------------
\subsubsection{Backtracking on the GPU}
The backtracking is processed similarly as with the Scala parser, the major difference being that since we are generating the C code, we can provide an immediate mapping from the subrule index to the backtrack elements to add to the trace. The backtrack is done in 3 steps:\ul
\item If window is set, the windowing aggregation kernel is run to determine the position of the best result within the matrix. Otherwise the best position can be found at the last computed element of the matrix.
\item For a $m\times n$ matrix, allocate a $m+n$ vector with two heads (reading, writing, initialized at the same position). Write the best element in the vector.
\item While there is a vector element that has been written but not read\ul
	\item From the parser id and its position retrieve the corresponding (subrule, concatenation indices) pair by reading in the corresponding matrix cell
	\item Using this, write new backtrack items that are at the origin of the current element.
	\ule
\ule

Since code is generated, it is possible to write the last step using a switch case, hereby flattening the writes in the vector (compared to recursive calls in Scala). Finally, since the trace has to be reversed, we can obtain this transformation for free by constructing the trace list from the end in the JNI conversion. Reversing the list present the advantage that trace is immediately usable to construct the desired element. It might be possible that Scala and CUDA parser provide different traces to construct the same result, because the trace verifies the dependency order, which is only a partial order.

% ----------------------------------------------
\subsubsection{User functions generation} \label{user_fun_gen}
{\color{red}
\begin{verbatim}
XXX: How LMS is used to generate function
Possibly write that we can do the same with macros
Why cannot use LMS for everything
Where and how is LMS useful
Also user function code generation with macros ?
\end{verbatim}
}

\subsection{Runtime execution engine}
The runtime execution engine is made of two instrumented compilers:\ul
\item A wrapper for {\tt g++} and {\tt nvcc} that can combine different file types ({\tt .h}, {\tt .c}, {\tt .cu}) into a JNI library which is then loaded into the current JVM instance. If necessary, paths can be customized to fit the user environment.
\item A wrapper for the Scala compiler, which allow the creation of Scala interface to the freshly compiled JNI libraries. It should be noted there that using {\tt VirtualDirectory} as compilation target prevents the interaction with JNI, hence physical path has to be used.
\ule

These two compilers interfaces are then mixed in another class that transform the previously (see \ref{codegen}) generated code, fixing input sizes and splits (number of kernels to launch to respect the time limit) constants, and execute it.

{\center\color{red} XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\\ CONTINUE HERE\\ 
XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX}

\subsection{LibRNA (?)}

The project development was scheduled in \ul
\item Problems analysis
\item Ad-hoc implementation (CUDA)
\item Porting ADP parsers
\item Improving parsers / adding functionalities
\item Integrating CUDA generation into parsers
\ule

### 2013-01-05 docs/report/report.tex
NIts.
Nits.

\section{Future work} \ul
\item Serial for problems larger than memory, use hybrid (Myers and Miller's algorithm with $\log_k$ with $k$ depending on available memory) approach depending available memory
\item Annotation on recursive functions to use dynamic programming like \\
	{\tt @DynaProg def Fib(n:Int) = if (n<=2) return 1 else Fib(n-1)+Fib(n-2)}.
\ule

@@

Possible extensions / optimizations:
- making serializable by using aggregation functions/transformations
- avoiding the cost matrix by moving data in the wavefront
- Scheduling for large problems (with memory loads for non-serializable)
- add FPGA target platform
- pack the data => less memory transfer (i.e. GATC=>4 letters in 1 char)
- operate on some larger word (ex 64 bits) to increase thread locality and reduce memory accesses

Core function F:
- in: s,t strings, neighbor costs: top, left, top+left, neighbor stats: top, left, top+left
- out: backtrack information, cost(i,j), stats
        stats (x)
         ||
         vv
stats -> KK -> new_stats (y')
(y)      || \
         vv  --+ backtrack info (Bxy)
     new_stats (x')

For stats (aggregation of the table): we need to maintain both an horizontal and vertical front
For backtracking we need to maintain the whole table
... | . |
..A | B | min/max: Min(D)=Min(Min(C),Min(B))
----+---+ sum    : Sum(D)=Sum(C)+Sum(B)-Sum(A)
..C | D | avg    : sum both #cells and values then divide at appropriate cell
----+---+
we many not need to store what's the previous cell, however, the previous
cell information (usually limited range) is much more compact than the score(32-64 bits)
\end{verbatim}

@@

\subsubsection*{Todo @TCK}\ul
\item Test/proof parsers are correct --- make sure implementation is correct
\item Automate test to compare against implementation
\item Benchmarks -- use extensively CUDA profiler
\item Write report
\item Port LibRNA for CUDA?
\ule

@@

% ------------------------------------------------------------------------------------------------
\newpage
\bibliographystyle{plain}
\bibliography{bibliography.bib}
\end{document}

>>>>

\section{Future work} \ul
\item Implementation for serial problems larger than memory, using hybrid (Myers and Miller's algorithm with $\log_k$ with $k$ depending on available memory)
\item Transforming non-serial into serial using aggregation functions/transformations
\item Avoid to maintain the cost matrix in memory by moving data in the wavefront
\item Non-serial scheduling for large problems (with memory loads for non-serializable)
\item Add FPGA target platform (George Nithin)
\item Pack the data => less memory transfer (i.e. GATC=>4 letters in 1 char)
\item Operate on larger words (ex 64 bits) to increase thread locality and reduce memory accesses
\ule

@@

Core function F:
- in: s,t strings, neighbor costs: top, left, top+left, neighbor stats: top, left, top+left
- out: backtrack information, cost(i,j), stats
         stats (x)
          ||
          vv
stats -> [  ] -> new_stats (y')
(y)       || \
          vv  --+ backtrack info (Bxy)
      new_stats (x')

For stats, we need to maintain an horizontal and vertical front
For backtracking we need to maintain the whole table
... | . |
..A | B | min/max: Min(D)=Min(Min(C),Min(B))
----+---+ sum    : Sum(D)=Sum(C)+Sum(B)-Sum(A)
..C | D | avg    : sum both #cells and values then divide at appropriate cell
----+---+
we may not need to store what's the previous cell, however, the previous
cell information (usually limited range) is much more compact than the score(32-64 bits)
\end{verbatim}

@@

\subsubsection*{Todo @TCK}\ul
\item Test/proof parsers are correct -- make sure implementation is correct
\item Automate test to compare against implementation
\item Benchmarks -- use CUDA profiler(?)
\item Write report
\item Port LibRNA for CUDA?
\ule

@@

% ------------------------------------------------------------------------------------------------
\newpage
%\usepackage{multicol}
%\usepackage{etoolbox}
%\patchcmd{\thebibliography}{\section*{\refname}}{\begin{multicols}{2}[\section*{\refname}]}{}{}
%\patchcmd{\endthebibliography}{\endlist}{\endlist\end{multicols}}{}{}
%\bibliographystyle{acm}
\bibliographystyle{plain}
\bibliography{bibliography.bib}
\end{document}
